### 2022-03-24 documentation in progress/paper.tex
Documentation Update: Sources

\section{Multi-Agent Pathfinding}
In Multi-Agent Pathfinding problems, you have a set of nodes, agents (e.g. robots), start and goal positions, and possible moves. The set of nodes declares the possible positions where robots can be and move to. The robots can move from one node to accessible neighboring nodes in discrete time steps. In this paper, only orthogonal moves are considered. The goal is to let each robot find a way from its start to its goal position. It is possible to not have predetermined goal positions for each specific robot and instead let the solver assign the goal positions to the robots that can have the best way to that position. I only dealt with having a predetermined goal position for each robot in this project. The result of the solving process is a plan with a set of moves for each robot. When executing the plan there must not be any collisions, i.e. two robots going through each other or two robots moving to the same node (see figure \ref{fig:mapf}). The last time step of the plan is called horizon or makespan INSERT SOURCE FOR MAPF.

@@

\section{Answer Set Programming}
Answer Set Programming is a declarative programming language that is especially useful when using it for NP-hard problems. For a complete introduction refer to INSERT ASP SOURCE. In this section, I will only briefly explain the main parts relevant to my encoding. \\
An ASP program contains a set of rules and atoms. The result is a stable model that is built upon these rules and atoms. An atom has the form: 
\begin{verbatim}
a.
\end{verbatim}
This means that the $a$ is treated as a fact and always has to be true. Rules have the form:
\begin{verbatim}
b :- a.
\end{verbatim}
which means that if $a$ is true then $b$ also has to be true. There are also negations:
\begin{verbatim}
b :- not a.
\end{verbatim}
This means that $b$ can only be inferred if $a$ is never true in the model. There are also choice rules:
\begin{verbatim}
x {b,c,d} y :- a.
\end{verbatim}
The atom $a$ is used to infer the choice rule. The solver can then decide which of the atoms $b$, $c$, and $d$ he wants to infer. He can also decide to not choose any atoms at all. You can set bounds for the number of atoms chosen. In this case, it means that the solver has to at least choose $x$ atoms, and cannot choose more than $y$ atoms. You can also influence the chosen atoms by using constraint rules of the form:
\begin{verbatim}
:- a, b.
\end{verbatim}
This means that $a$, and $b$ cannot be true at the same time. Combining this constraint rule with the choice rule above means that the solver will not choose to infer $b$. Atoms can also have attributes:
\begin{verbatim}
a(1). a(2).
\end{verbatim}
This means that there is an atom $a$ with the attribute $1$, and an atom $a$ with the attribute $2$. In rules you can exchange the attributes with variables:
\begin{verbatim}
b(X) : a(X).
\end{verbatim}
This means that the solver looks for every atom $a$ with exactly one attribute and infers the atom $b$ with the same attribute. Before the solver looks for stable models, the program is $grounded$. During this process, the variables in the program are exchanged with the actual values. \\
The last form of rule to talk about is called aggregates. I will explain two aggregates that have also been used in the abstraction encoding. The first one is $min$:
\begin{verbatim}
a(MIN) :- MIN == #min{X: b(X)}.
\end{verbatim}
The solver looks for the attribute of the atom $b$ with the lowest value and assigns  $MIN$ this lowest value. It then infers the atom $a$ with the attribute $MIN$. The other aggregate, I used in my encoding is $minimize$:
\begin{verbatim}
#minimize {X : a(X)}.
\end{verbatim}
This means that the solver looks at every atom $a$ with one attribute and builds the sum of these attributes. It then looks for the model with the lowest sum. \\
There are also other aggregates that I have not explained but since I do not use them in my encoding, I will not explain them. An ASP program has the file ending $.lp$.\\
To ground and solve the programs in this paper, clingo-5.4.0 was used. INSERT SOURCE CLINGO

@@

\subsection{Node Combining}
The idea for the node combining was taken from two other papers looking for abstractions (CITE PAPERS). The approach had to be slightly altered because my goal is to only build an abstraction and not solve the instance completely. The idea is two combine several nodes together into one node (see figure \ref{post-com} for illustration). Afterward, the solver looks for the shortest path in the combined instance. The solver then outputs all the nodes that have been visited in the combined form (see figure \ref{fig:onc-result}. The combining of nodes looks like this:
\begin{verbatim}
comb_position((X,Y),(X',Y'))   :- position((X,Y)),
          X' = (X+(size_x-1))/size_x, Y' = (Y+(size_y-1))/size_y.
comb_position(C)               :- comb_position(_,C).
comb_position(R,C',0)          :- position(R,C,0), comb_position(C,C').
comb_position(shelf(S),C',0)   :- position(shelf(S),C,0), comb_position(C,C').
\end{verbatim}
The first rule calculates the coordinates that a node has after combining. The constants $size\_x$ and $size\_y$ say how many nodes in x- and y-direction are combined together. The second rule is used so that it is not necessary to always care about the initial nodes. Rule 3 and 4 change the positions of robots and shelves to their new coordinates. The rest of the encoding is only changed so that the atoms $comb\_position$ instead of $position$ are used. The output atom $imp\_position$ is then easily inferred by looking at the positions visited by robots:
\begin{verbatim}
imp_position(R,C) :- isRobot(robot(R)), comb_position(R,C',_), comb_position(C,C').
\end{verbatim}
It takes the positions visited by robots, looks at the nodes that were combined into that position, and then infers the respective atoms. \\
Simply using this approach has the problem that walls are completely ignored while building the combined instance. This means that the encoding introduced so far can only be used for open instances, i.e. instances that do not have any walls except for the outer walls. This is why I call this approach ``Open Node Combining''. The advantage of this method is that the combining size can be freely chosen, as long as positive integers are used. \\
To also use the node combining approach in instances with walls, I created the approach ``Complete Node Combining''. This approach mostly works the same but it can only be used when combining 2x2 squares together. The difference is that a $wall$-atom is introduced so that walls are considered after the combining:
\begin{verbatim}
wall((X1',Y1'),(X2',Y2')) :- comb_position((X1',Y1')), x_arr(X1), y_arr(Y1),
    not position((X1,Y1)), not position((X1+1,Y1)),
    X1' = (X1+(size_x-1))/size_x, Y1' = (Y1+(size_y-1))/size_y,
    X1'-1 = X1/size_x, Y1'-1 = Y1/size_y,
    comb_position((X2',Y2')), X1'=X2', Y1'-1=Y2'.
\end{verbatim}
This is one of four rules inferring a $wall$-atom. The idea is that in a 2x2 square there are four possible walls: two nodes missing to the top, bottom, left, or right. Each rule inferring a $wall$-atom is used for one of the possible walls. The rule takes a combined node and checks if there are exactly two uncombined nodes missing that would be in that combined node if they existed. The $nextto$-atom is then extended to also check if a wall is between two combined nodes before inferring the respective atom:
\begin{verbatim}
nextto((X,Y),(DX,DY),(X',Y')) :- dir((DX,DY)),
   comb_position((X,Y)), comb_position((X',Y')),
   (X,Y)=(X'-DX,Y'-DY), (X',Y')=(X+DX,Y+DY),
   not wall((X,Y),(X',Y')), not wall((X',Y'),(X,Y)).
\end{verbatim}
As mentioned before, this way of introducing walls leads to only being able to combine 2x2 squares. I did not find a way to combine bigger sizes because looking at walls in these cases gets much more complicated. \\

@@

\subsection{Reachable Nodes}
The last abstraction method was also inspired by a research paper (CITE PAPER). Again, a slight alteration was done because the goal is to only find an abstraction and not a complete solution. The idea is to find the shortest path for each robot as always. Then, the solver looks at how many steps each node close to the path of the robots is deviating from that path. The path itself is deviating by zero steps. Each node next to the path is deviating by one step, the nodes right next to them by 2  steps, and so on (see figure \ref{fig:rn}). The maximum amount of steps looked at is the number of steps the robot needed in total to reach its goal. 
\begin{verbatim}
makespan(R,N) :- move(R,_,N).
imp_pos(R,C,0) 	:- isRobot(robot(R)), position(R,C,_).
imp_pos(R,C,N) 	:- isRobot(robot(R)), imp_pos(R,C',N-1), makespan(R,N), nextto(C',_,C).
imp_position(R,C,MIN) 		:- imp_pos(R,C,MIN), MIN == #min{X: imp_pos(R,C,X)}.
\end{verbatim}
The first line looks at how many steps each robot takes. It is important to note that makespan refers to an individual robot whereas horizon refers to all robots. This means that the highest makespan is equal to the horizon. Another point to note is that the $makespan$-atom not only has the highest time step for an individual robot but each step it takes infers a $makespan$-atom. The second line initializes the path of the robot itself as deviating by zero steps. The third line then infers the next set of positions and the respective deviation until the makespan of the robot is reached. This way of inferring these atoms leads to having multiple $imp\_pos$-atoms for a single node. The last line makes sure that only the smallest deviation gets taken for the $imp\_position$-atoms. This time the $imp\_position$-atoms, as well as the $new\_init$-atoms, have three attributes because the amount of deviating steps is important to output. The abstraction this time does not simply consist of a set of nodes but instead multiple sets of nodes with a different distance to the shortest path. A solver using this abstraction could then decide how big the distance should be and only take a subset of the nodes provided to solve an instance. \\
There was also a slightly different approach that did not look at all the nodes up to the complete makespan of a robot. Instead, it only looked at how many steps were still left to reach the goal and calculated the deviation up to these remaining steps. This method provides less information but the hope was that it would find the solution faster. During the functionality testing, it already became clear that the solving time with this different approach was slower than the original approach. This is why this different approach was discarded during the creation process.

@@

The benchmarking was done on the Turing-Cluster of the university of Potsdam (LINK TO HPC SITE). A single run was calculated in sequential form. Each node of the cluster has 2 CPUs of the type Intel Xeon E5-2650v4, 64 GB of RAM, and an FDR Infiniband Adapter. All results can be found in the appendix.

@@

There are six map layouts used for the benchmarking. Each method was tested 10 times with preset starting coordinates and 10 times with random starting coordinates on each map. Some of the layouts were inspired by another paper (CITE BENCHMARKING PAPER). Three layouts are just an open map, which means that they have no inner walls. They have the sizes 16x16 nodes, 64x64 nodes, and 128x128 nodes. The preset coordinates for this type of map were chosen in a way that leads to at least one robot having the maximum amount of distance possible to its goal. The robots are added from left to right and top to bottom while the shelves are added the other way round. \\

@@

\begin{thebibliography} {3}
\bibitem{my-git}
My Github repository
\begin{verbatim}
https://github.com/salewsky/MAPF-Project
\end{verbatim}

\bibitem{tarek-git}
Github repository of Tarek Ramadan
\begin{verbatim}
https://github.com/tramadan-up/mapf-ba
\end{verbatim}

\bibitem{asprilo}
Asprilo website
\begin{verbatim}
https://potassco.org/asprilo/
\end{verbatim}

%@article{stern2019mapf,
  %title={Multi-Agent Pathfinding: Definitions, Variants, and Benchmarks},
  %author={Roni Stern and Nathan R. Sturtevant and Ariel Felner and Sven Koenig and Hang Ma and Thayne T. Walker and Jiaoyang Li and Dor Atzmon and Liron Cohen and T. K. Satish Kumar and Eli Boyarski and Roman Bartak},
 % journal={Symposium on Combinatorial Search (SoCS)},
 % year={2019},
 % pages={151--158}
%}

>>>>

\section{Multi-Agent Pathfinding}
In Multi-Agent Pathfinding problems, you have a set of nodes, agents (e.g. robots), start and goal positions, and possible moves. The set of nodes declares the possible positions where robots can be and move to. The robots can move from one node to accessible neighboring nodes in discrete time steps. In this paper, only orthogonal moves are considered. The goal is to let each robot find a way from its start to its goal position. It is possible to not have predetermined goal positions for each specific robot and instead let the solver assign the goal positions to the robots that can have the best way to that position. I only dealt with having a predetermined goal position for each robot in this project. The result of the solving process is a plan with a set of moves for each robot. When executing the plan there must not be any collisions, i.e. two robots going through each other or two robots moving to the same node (see figure \ref{fig:mapf}). The last time step of the plan is called horizon or makespan \cite{mapf}.

@@

\section{Answer Set Programming}
Answer Set Programming is a declarative programming language that is especially useful when using it for NP-hard problems. For a complete introduction refer to \cite{asp}. In this section, I will only briefly explain the main parts relevant to my encoding. \\
An ASP program contains a set of rules and atoms. The result is a stable model that is built upon these rules and atoms. An atom has the form: 
\begin{verbatim}
a.
\end{verbatim}
This means that the $a$ is treated as a fact and always has to be true. Rules have the form:
\begin{verbatim}
b :- a.
\end{verbatim}
which means that if $a$ is true then $b$ also has to be true. There are also negations:
\begin{verbatim}
b :- not a.
\end{verbatim}
This means that $b$ can only be inferred if $a$ is never true in the model. There are also choice rules:
\begin{verbatim}
x {b,c,d} y :- a.
\end{verbatim}
The atom $a$ is used to infer the choice rule. The solver can then decide which of the atoms $b$, $c$, and $d$ he wants to infer. He can also decide to not choose any atoms at all. You can set bounds for the number of atoms chosen. In this case, it means that the solver has to at least choose $x$ atoms, and cannot choose more than $y$ atoms. You can also influence the chosen atoms by using constraint rules of the form:
\begin{verbatim}
:- a, b.
\end{verbatim}
This means that $a$, and $b$ cannot be true at the same time. Combining this constraint rule with the choice rule above means that the solver will not choose to infer $b$. Atoms can also have attributes:
\begin{verbatim}
a(1). a(2).
\end{verbatim}
This means that there is an atom $a$ with the attribute $1$, and an atom $a$ with the attribute $2$. In rules you can exchange the attributes with variables:
\begin{verbatim}
b(X) : a(X).
\end{verbatim}
This means that the solver looks for every atom $a$ with exactly one attribute and infers the atom $b$ with the same attribute. Before the solver looks for stable models, the program is $grounded$. During this process, the variables in the program are exchanged with the actual values. \\
The last form of rule to talk about is called aggregates. I will explain two aggregates that have also been used in the abstraction encoding. The first one is $min$:
\begin{verbatim}
a(MIN) :- MIN == #min{X: b(X)}.
\end{verbatim}
The solver looks for the attribute of the atom $b$ with the lowest value and assigns  $MIN$ this lowest value. It then infers the atom $a$ with the attribute $MIN$. The other aggregate, I used in my encoding is $minimize$:
\begin{verbatim}
#minimize {X : a(X)}.
\end{verbatim}
This means that the solver looks at every atom $a$ with one attribute and builds the sum of these attributes. It then looks for the model with the lowest sum. \\
There are also other aggregates that I have not explained but since I do not use them in my encoding, I will not explain them. An ASP program has the file ending $.lp$.\\
To ground and solve the programs in this paper, clingo-5.4.0 was used \cite{clingo}.

@@

\subsection{Node Combining}
The idea for the node combining was taken from two other papers looking for abstractions \cite{nc1} \cite{nc2}. The approach had to be slightly altered because my goal is to only build an abstraction and not solve the instance completely. The idea is two combine several nodes together into one node (see figure \ref{post-com} for illustration). Afterward, the solver looks for the shortest path in the combined instance. The solver then outputs all the nodes that have been visited in the combined form (see figure \ref{fig:onc-result}. The combining of nodes looks like this:
\begin{verbatim}
comb_position((X,Y),(X',Y'))   :- position((X,Y)),
          X' = (X+(size_x-1))/size_x, Y' = (Y+(size_y-1))/size_y.
comb_position(C)               :- comb_position(_,C).
comb_position(R,C',0)          :- position(R,C,0), comb_position(C,C').
comb_position(shelf(S),C',0)   :- position(shelf(S),C,0), comb_position(C,C').
\end{verbatim}
The first rule calculates the coordinates that a node has after combining. The constants $size\_x$ and $size\_y$ say how many nodes in x- and y-direction are combined together. The second rule is used so that it is not necessary to always care about the initial nodes. Rule 3 and 4 change the positions of robots and shelves to their new coordinates. The rest of the encoding is only changed so that the atoms $comb\_position$ instead of $position$ are used. The output atom $imp\_position$ is then easily inferred by looking at the positions visited by robots:
\begin{verbatim}
imp_position(R,C) :- isRobot(robot(R)), comb_position(R,C',_), comb_position(C,C').
\end{verbatim}
It takes the positions visited by robots, looks at the nodes that were combined into that position, and then infers the respective atoms. \\
Simply using this approach has the problem that walls are completely ignored while building the combined instance. This means that the encoding introduced so far can only be used for open instances, i.e. instances that do not have any walls except for the outer walls. This is why I call this approach ``Open Node Combining''. The advantage of this method is that the combining size can be freely chosen, as long as positive integers are used. \\
To also use the node combining approach in instances with walls, I created the approach ``Complete Node Combining''. This approach mostly works the same but it can only be used when combining 2x2 squares together. The difference is that a $wall$-atom is introduced so that walls are considered after the combining:
\begin{verbatim}
wall((X1',Y1'),(X2',Y2')) :- comb_position((X1',Y1')), x_arr(X1), y_arr(Y1),
    not position((X1,Y1)), not position((X1+1,Y1)),
    X1' = (X1+(size_x-1))/size_x, Y1' = (Y1+(size_y-1))/size_y,
    X1'-1 = X1/size_x, Y1'-1 = Y1/size_y,
    comb_position((X2',Y2')), X1'=X2', Y1'-1=Y2'.
\end{verbatim}
This is one of four rules inferring a $wall$-atom. The idea is that in a 2x2 square there are four possible walls: two nodes missing to the top, bottom, left, or right. Each rule inferring a $wall$-atom is used for one of the possible walls. The rule takes a combined node and checks if there are exactly two uncombined nodes missing that would be in that combined node if they existed. The $nextto$-atom is then extended to also check if a wall is between two combined nodes before inferring the respective atom:
\begin{verbatim}
nextto((X,Y),(DX,DY),(X',Y')) :- dir((DX,DY)),
   comb_position((X,Y)), comb_position((X',Y')),
   (X,Y)=(X'-DX,Y'-DY), (X',Y')=(X+DX,Y+DY),
   not wall((X,Y),(X',Y')), not wall((X',Y'),(X,Y)).
\end{verbatim}
As mentioned before, this way of introducing walls leads to only being able to combine 2x2 squares. I did not find a way to combine bigger sizes because looking at walls in these cases gets much more complicated. \\

@@

\subsection{Reachable Nodes}
The last abstraction method was also inspired by a research paper \cite{rn}. Again, a slight alteration was done because the goal is to only find an abstraction and not a complete solution. The idea is to find the shortest path for each robot as always. Then, the solver looks at how many steps each node close to the path of the robots is deviating from that path. The path itself is deviating by zero steps. Each node next to the path is deviating by one step, the nodes right next to them by 2  steps, and so on (see figure \ref{fig:rn}). The maximum amount of steps looked at is the number of steps the robot needed in total to reach its goal. 
\begin{verbatim}
makespan(R,N) :- move(R,_,N).
imp_pos(R,C,0) 	:- isRobot(robot(R)), position(R,C,_).
imp_pos(R,C,N) 	:- isRobot(robot(R)), imp_pos(R,C',N-1), makespan(R,N), nextto(C',_,C).
imp_position(R,C,MIN) 		:- imp_pos(R,C,MIN), MIN == #min{X: imp_pos(R,C,X)}.
\end{verbatim}
The first line looks at how many steps each robot takes. It is important to note that makespan refers to an individual robot whereas horizon refers to all robots. This means that the highest makespan is equal to the horizon. Another point to note is that the $makespan$-atom not only has the highest time step for an individual robot but each step it takes infers a $makespan$-atom. The second line initializes the path of the robot itself as deviating by zero steps. The third line then infers the next set of positions and the respective deviation until the makespan of the robot is reached. This way of inferring these atoms leads to having multiple $imp\_pos$-atoms for a single node. The last line makes sure that only the smallest deviation gets taken for the $imp\_position$-atoms. This time the $imp\_position$-atoms, as well as the $new\_init$-atoms, have three attributes because the amount of deviating steps is important to output. The abstraction this time does not simply consist of a set of nodes but instead multiple sets of nodes with a different distance to the shortest path. A solver using this abstraction could then decide how big the distance should be and only take a subset of the nodes provided to solve an instance. \\
There was also a slightly different approach that did not look at all the nodes up to the complete makespan of a robot. Instead, it only looked at how many steps were still left to reach the goal and calculated the deviation up to these remaining steps. This method provides less information but the hope was that it would find the solution faster. During the functionality testing, it already became clear that the solving time with this different approach was slower than the original approach. This is why this different approach was discarded during the creation process.

@@

The benchmarking was done on the Turing-Cluster of the university of Potsdam \cite{hpc}. A single run was calculated in sequential form. Each node of the cluster has 2 CPUs of the type Intel Xeon E5-2650v4, 64 GB of RAM, and an FDR Infiniband Adapter. All results can be found in the appendix.

@@

There are six map layouts used for the benchmarking. Each method was tested 10 times with preset starting coordinates and 10 times with random starting coordinates on each map. Some of the layouts were inspired by another paper \cite{mapf}. Three layouts are just an open map, which means that they have no inner walls. They have the sizes 16x16 nodes, 64x64 nodes, and 128x128 nodes. The preset coordinates for this type of map were chosen in a way that leads to at least one robot having the maximum amount of distance possible to its goal. The robots are added from left to right and top to bottom while the shelves are added the other way round. \\

@@

\begin{thebibliography} {3}
\bibitem{my-git}
\begin{verbatim}
https://github.com/salewsky/MAPF-Project
\end{verbatim}

\bibitem{tarek-git}
\begin{verbatim}
https://github.com/tramadan-up/mapf-ba
\end{verbatim}

\bibitem{asprilo}
\begin{verbatim}
https://potassco.org/asprilo/
\end{verbatim}

\bibitem{mapf}
Stern et al. (2019). Multi-Agent Pathfinding: Definitions, Variants, and Benchmarks. Symposium on Combinatorial Search (SoCS), 151-158.

\bibitem{asp}
Janhunen, T., and Niemelä, I. (2016). The Answer Set Programming Paradigm. AI Magazine, 13-24.

\bibitem{clingo}
Gebser et al. (2010). A User's Guide to gringo, clasp, clingo, and iclingo.

\bibitem{nc1}
Sturtevant, N., and Buro, M. (2005). Partial Pathfinding Using Map Abstraction and Refinement. Proceedings of AAAI, 47-52.

\bibitem{nc2}
Sturtevant, Nathan and Buro, Michael. (2021). Improving Collaborative Pathfinding Using Map Abstraction. Proceedings of the AAAI Conference on Artificial Intelligence and Interactive Digital Entertainment, 2(1), 80-85.

\bibitem{rn}
Anonymous Author(s). Reduction-based Solving of Multi-agent Pathfinding on Large Maps Using Graph Pruning.

\bibitem{hpc}
\begin{verbatim}
https://www.cs.uni-potsdam.de/bs/research/labs.html#turing
\end{verbatim}
